---
title: "coxregress"
author: "josiah"
date: "9/20/2017"
output: html_document
---

```{r setup, collapse=TRUE, warning=FALSE}
library(survival)
library(survminer)
#load data
d0 = read.csv('/Users/josiah/Box Sync/projects/cuedti/data/ainsnaccvlpfc_meanmetric_tpincr_wide_170919.csv')
d1 = read.csv('/Users/josiah/Box Sync/projects/cuedti/data/relapse_data_170920_nacc.csv')
#remove subjects without relapse data
d1 = d1[is.na(d1$relapse) == FALSE,]
#make relapse a factor
d1$relapse = as.factor(d1$relapse)
colnames(d1)[1] = 'subject'

#create survival variable
d1$survival <- with(d1, Surv(obstime, censored == 0))
d1$survival

#merge relapse data and tract data
df = merge.data.frame(d1, d0, by = 'subject')
```

#cox regressions

ains-vlpfc tract MD has significant hazard ratio for relapse. </br>
(weaker tract predicts relapse) </br>
holds after controlling for age.

```{r ainsvlpfc, collapse=TRUE}
cox1 <- coxph(survival ~ vlpfcains_r_md_meantp50 + age, data =  df)
summary(cox1)
ggsurvplot(survfit(cox1), palette = "#2E9FDF",ggtheme = theme_minimal())
```

ains-nacc tract FA does not itself have a significant hazard ratio for relapse.

```{r ainsnacc, collapse=TRUE}
cox2 <- coxph(survival ~ ainsnacc_r_fa_meanincr10 + age, data =  df)
summary(cox2)
```

both tracts in same model have significant or trending hazard ratio for relapse, and in the right directions.

```{r both, collapse=TRUE}
cox3 <- coxph(survival ~ vlpfcains_r_md_meantp50 + ainsnacc_r_fa_meanincr10, data =  df)
summary(cox3)
```

controlling for age, only ains-nacc tract MD is significant.

```{r both age, collapse=TRUE}
cox4 <- coxph(survival ~ vlpfcains_r_md_meantp50 + ainsnacc_r_fa_meanincr10 + age, data =  df)
summary(cox4)
```

#pairs plots

```{r pairs plot, collapse=TRUE, warning=FALSE}
require(psych)
dp = with(df, data.frame(relapse, obstime, age, vlpfcains_r_md_meantp50, ainsnacc_r_fa_meanincr10))
pairs.panels(dp)
```

Weird distributions for age and ains-nacc FA. </br>
One subject's ains-nacc tract has leverage. </br>
One subject's time-to-relapse has laverage. </br>

Next: look at the high leverage subjects, and recover subjects who couldn't find track in first pass.


