---
title: "pairwise plots"
author: "Kelly MacNiven"
date: "4/3/2018"
output: html_document
---

## packages
```{r}
library(psych) 
library(lattice)
library(ggplot2)
library(knitr) # the package that renders R markdown and has some good additional functionality
```

# load data & set up
```{r getdata}
d = read.csv('/Users/kelly/cueexp/data/relapse_data/relapse_data_180330.csv')

# remove subjects without relapse data
d = d[is.na(d$relapse) == FALSE,]

# make relapse & subjects factors
d$relIn6Mos = as.factor(d$relIn6Mos)
d$relapse <- as.factor(d$relapse)
d$relIn6Mos <- as.factor(d$relIn6Mos)
d$Subject <- as.factor(d$subj)

# check out the data 
#kable(d)
```

### run paired plots 


#Descriptive stats
#```{r}
#summary(d)  #print out the min, max, range, mean, median, etc. of the data
#describe(d)
#```

## plots
```{r pairplots}
#draws scatterplots, histograms, and shows correlations  (source found in the psych package)
png(filename = '/Users/kelly/cueexp/figures/relapse_prediction/pairplots.png',
     units="in",
     width=5,
     height=5,
     res=300)
pairs.panels(d)
dev.off()
```

# ```{r}
# #my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")  
# my_cols <- c("#00AFBB", "#FC4E07")  
# pairs(dd, 
#       pch = 19,  
#       cex = 0.5,
#       col = my_cols[d$relapse],
#       lower.panel=NULL)
# 
# pairs.panels(dd,
#              hist.col = my_cols[2]
# )
# 
# # Correlation panel
# panel.cor <- function(x, y){
#   usr <- par("usr"); on.exit(par(usr))
#   par(usr = c(0, 1, 0, 1))
#   r <- round(cor(x, y), digits=2)
#   txt <- paste0("R = ", r)
#   cex.cor <- 0.8/strwidth(txt)
#   text(0.5, 0.5, txt, cex = cex.cor * r)
# }
# 
# # Customize upper panel
# upper.panel<-function(x, y){
#   points(x,y, pch = 19, col = my_cols[as.factor(d$relapse)])
# }
# # Create the plots
# pairs(dd, 
#       lower.panel = panel.cor,
#       upper.panel = upper.panel)
# ```
# 
# 
# pairs.panels(attitude)   #see the graphics window
# data(iris)
# pairs.panels(iris[1:4],bg=c("red","yellow","blue")[iris$Species],
#              pch=21,main="Fisher Iris data by Species") #to show color grouping
# 
# pairs.panels(iris[1:4],bg=c("red","yellow","blue")[iris$Species],
#              pch=21+as.numeric(iris$Species),main="Fisher Iris data by Species",hist.col="red") 
# #to show changing the diagonal
# 
# #to show 'significance'
# pairs.panels(iris[1:4],bg=c("red","yellow","blue")[iris$Species],
#              pch=21+as.numeric(iris$Species),main="Fisher Iris data by Species",hist.col="red",stars=TRUE) 
# 
# 
# 
# #demonstrate not showing the data points
# data(sat.act)
# pairs.panels(sat.act,show.points=FALSE)
# #better yet is to show the points as a period
# pairs.panels(sat.act,pch=".")
# #show many variables with 0 gap between scatterplots
# # data(bfi)
# # pairs.panels(bfi,show.points=FALSE,gap=0)
# 
# #plot raw data points and then the weighted correlations.
# #output from statsBy
# sb <- statsBy(sat.act,"education")
# pairs.panels(sb$mean,wt=sb$n)  #report the weighted correlations
# #compare with 
# pairs.panels(sb$mean) #unweighed correlations
# 
# 
# 
# 
# 
